Constraining the Interiors of Asteroids Through Close Encounters

### 2022-03-28 paper/main.tex
Staged new density moments

Given the small effect of $\epsilon$ on $K_{\ell m}$, it might be tempting to neglect the effect of planetary oblateness when fitting $K_{\ell m}$ to data. However, the bottom pane of figure \ref{fig:scan-oblateness} demonstrates that this is not valid. This figure displays $K_{\ell m}$ as extracted by a fit assuming $\epsilon = 0$, but run on data generated by assuming nonzero $\epsilon$. The difference between the posterior mean $K_{\ell m}$ and true $K_{\ell m}$ are shown. Posterior uncertainties are also shown as bands. The figure shows that even for low (Moon-scale) oblateness, the fit results are inconsistent with the true $K_{\ell m}$ values, since $|\Delta K_{\ell m} = 0$ is not contained in the 2-$\sigma$ band. This effect is much worse for large oblateness, growing to a difference on the order of $10^{-2}$ between the true and fit parameters for large oblateness. Therefore, accurately modeling oblateness to high precision is essential for accurate estimation of fit parameters. For non-equatorial orbits, with $J_{22} \neq 0$, we also expect $J_{22}$ to affect the accuracy of the fit results to a similar degree, so it is also vital to use the correct asteroid orbital plane.

@@

First, we simulate several asteroids with uniform density distributions and ask whether a distribution is recovered. The shapes we use are the symmetric and asymmetric reference ellipsoids, a ``dumbbell'' (two spheres of equal radius conjoined such that the surface of one intersects the center of the other), and a tetrahedron. The extracted distributions for all three models are shown in figure \ref{fig:uniform-density}. Also in each figure is $\Delta K_{\ell m}$, the magnitude of difference between the fitted density moments and the density moments of the final distribution. This difference is nonzero due to numerical error in the likelihood model case, but for the harmonic and the lumpy models (which are not guaranteed to exactly reproduce the input density moments), they express the inability of the density model to exactly reproduce the density moments.
\begin{figure*}
  \centering
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-harmonic.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-lumpy.pdf}

@@

The lumpy model produces zero-mass lumps in this case, and produces very small $\Delta K_{\ell m}$. \jtd{What are the exact parameters of the lump? Look at log files}.

@@

We also test several non-uniform density asteroids and compare the generated density distributions for all three models to see if the non-uniformities are recovered. We consider two asteroids, one with mass concentrated in the middle, and one with mass concentrated symmetrically around the edges. (Specifically, the mass distribution follows a Gaussian which falls to a fraction $e^{\pm 0.5}$ of the central density on the surface of the asteroid.) To prevent the problems seen in previous sections when a symmetric asteroid is simulated, we set the shape of the asteroids to the asymmetric reference asteroid. The extracted density distributions are shown in figure \ref{fig:non-uniform-density}

@@

\jtd{Why do the densities go negative?}

\jtd{What's wrong with the lumpy model? Wasn't it giving really good results before?}

@@

However, the lumpy model exactly computes the location and density of the mass. \jtd{But it's not quite exact. Why?}. \jtd{Would like to say that there's no error in klm.} We see therefore that the lumpy model is capable of recreating the location of lumps accurately (in simple cases such as this). But the ability of the lumpy model to place a lump does not guarantee the lump to actually exist, as was seen in all the other asteroid distributions analyzed here.

@@

We warmly thank Emmanuel Jehin, Maxime Devogele, and Marin \jtd{What's Marin's last name?} for meeting with the authors to discuss this work, and pointing out possible future directions to take it. JTD would like to thank the Massachusetts Institute of Technology Undergraduate Research Opportunities Program (MIT UROP) office for funding his work for many semesters, and their flexibility in funding students from departments other than the source of funding. The work of JdW was funded by \textit{Who funds Julien?}. Much of the computation shown in this paper was executed on MIT Supercloud's facilities.

@@

\begin{enumerate}
  \item An orbit around Earth with $6$ km s$^{-1}$ excess velocity and perigee at 5 Earth radii. This orbit was chosen to roughly match that of \jtd{Number} Apophi s\jtd{Cite Apophis close approach}
  \item An initial roll of $\gamma_0=\pi/8$.
  \item A cadence of 2 minutes and observational uncertainty of $\sigma_\theta = 0.01$ and $\sigma_\rho / \sigma_\theta = 10^{-5}$.
  \item A rotational period of 9 hours, with the rotational velocity vector distributed between the $\unit X$, $\unit Y$, and $\unit Z$ axes in a $1:2:-2$ ratio.
  \item An asteroid with radius $a_m = 1$ km and $K_{3m}=0$. For $K_{22}$ and $K_{20}$, we use two standard values. One with $(K_{22}, K_{20}) = (0, -0.097)$ and one with $(0.052, -0.202)$. Including the third point obtained by reflection $K_{22}\rightarrow -K_{22}$, these are the three points that minimize the mean distance between an arbitrary point in the allowed parameter space (equation \ref{eqn:parameter-bounds}) and these reference values. The first point is called the symmetric case because the corresponding uniform-density-ellipsoid model is rotationally symmetric around $\unit z$. The second case and its reflection are called the asymmetric cases. Values of $(0.052, -0.202)$ have $a < b$ in the ellipsoid model, and the reflected value has $a > b$. If not specified, we use the $a < b$ case. Specifically, the asymmetric case has $a=1140$ m, $b=1839$ m, and $c=565$ m, while the symmetric case has $a=b=1411$ m and $c=1008$ m.
\end{enumerate}

>>>>

Given the small effect of $\epsilon$ on $K_{\ell m}$, it might be tempting to neglect the effect of planetary oblateness when fitting $K_{\ell m}$ to data. However, the bottom pane of figure \ref{fig:scan-oblateness} demonstrates that this is not valid. This figure displays $K_{\ell m}$ as extracted by a fit assuming $\epsilon = 0$, but run on data generated by assuming nonzero $\epsilon$. The difference between the posterior mean $K_{\ell m}$ and true $K_{\ell m}$ are shown. Posterior uncertainties are also shown as bands. The figure shows that even for low (Moon-scale) oblateness, the fit results are inconsistent with the true $K_{\ell m}$ values, since $|\Delta K_{\ell m}| = 0$ is not contained in the 2-$\sigma$ band. This effect is much worse for large oblateness, growing to a difference on the order of $10^{-2}$ between the true and fit parameters for large oblateness. Therefore, accurately modeling oblateness to high precision is essential for accurate estimation of fit parameters. For non-equatorial orbits, with $J_{22} \neq 0$, we also expect $J_{22}$ to affect the accuracy of the fit results to a similar degree, so it is also vital to use the correct asteroid orbital plane.

@@

First, we simulate several asteroids with uniform density distributions and ask whether a distribution is recovered. The shapes we use are the symmetric and asymmetric reference ellipsoids, a ``dumbbell'' (two spheres of equal radius conjoined such that the surface of one intersects the center of the other), and a tetrahedron. The extracted distributions for all three models are shown in figure \ref{fig:uniform-density}. Also in each figure is $\Delta K_{\ell m}$, the squared magnitude of difference between the fitted density moments and the density moments of the final distribution. This difference is nonzero due to numerical error in the likelihood model case, but for the harmonic and the lumpy models (which are not guaranteed to exactly reproduce the input density moments), they express the inability of the density model to exactly reproduce the density moments.
\begin{figure*}
  \centering
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-likelihood.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-harmonic.pdf}\hfill
  \includegraphics[width=0.33\textwidth]{figs/asym-ell-lumpy.pdf}

@@

The lumpy model produces lumps so large that they dominate the asteroid, predicting uniform distribution with $K_{\ell m} = 0$ for $\ell \geq 1$.. The error in density moments $\Delta K_{\ell m}$ is therefore very large, equal to the sum of $|K_{\ell m}|$ for the original asteroids.

@@

We also test several non-uniform density asteroids and compare the generated density distributions for all three models to see if the non-uniformities are recovered. We consider two asteroids, one with mass concentrated in the middle, and one with mass concentrated around the edges. (Specifically, the mass distribution follows a spherically symmetric exponential $\rho(r) \propto e^{\pm r^2/a_m^2}$). To prevent the problems seen in previous sections when a symmetric asteroid is simulated, we set the shape of the asteroids to the asymmetric reference asteroid. The extracted density distributions are shown in figure \ref{fig:non-uniform-density}

@@

\jtd{Rewrite after new model comes through}

The lumpy model has one caveat that has not yet been illustrated, which applies for $N=d=1$ if (1) the center of mass of the asteroid is the centroid of the asteroid shape and (2) the density moments of the asteroid shape assuming uniform density are proportional to the fitted density moments. (1) guarantees that the lump will be placed at the origin and therefore will not contribute to $K_{\ell m}$ for $\ell \geq 1$, and (2) guarantees that the mass of the asteroid shape will be chosen such that the shape density moments exactly match the fitted density moments. The lump radius will then be set to zero because the fitted density moments are already matched. This zero-radius-lump result will occur even if the true density distribution has a large-radius lump at the origin.

An example of this caveat applied is an elliptical lump inside a similar elliptical asteroid. This is why, in the center-weighted and edge-weighted density distributions, we specifically chose a spherically symmetric density distribution for our ellipsoidal asteroid shape, rather than an ellipsoidal distribution. Thus, the caveat is avoided.

@@

However, the lumpy model exactly computes the location and density of the mass. \jtd{But it's not quite exact. Why?}. \jtd{Would like to say that there's no error in klm.} We see therefore that the lumpy model is capable of recreating the location of lumps accurately (except for the caveat mentioned in the previous section). But the ability of the lumpy model to place a lump does not guarantee the lump to actually exist, as was seen in all the other asteroid distributions analyzed here.

@@

We warmly thank Emmanuel Jehin, Maxime Devogele, and Marin \jtd{What's Marin's last name?} for meeting with the authors to discuss this work, and pointing out possible future directions to take it. JTD would like to thank the Massachusetts Institute of Technology Undergraduate Research Opportunities Program (MIT UROP) office for funding his work for many semesters, and their flexibility in funding students from departments other than the source of funding. The work of JdW was funded by \jtd{Who funds Julien?}. Much of the computation shown in this paper was executed on MIT Supercloud's facilities.

@@

\begin{enumerate}
  \item An orbit around Earth with $6$ km s$^{-1}$ excess velocity and perigee at 5 Earth radii. This orbit was chosen to roughly match that of 99942 Apophis \jtd{Cite Apophis close approach}
  \item An initial roll of $\gamma_0=\pi/8$.
  \item A cadence of 2 minutes and observational uncertainty of $\sigma_\theta = 0.01$ and $\sigma_\rho / \sigma_\theta = 10^{-5}$.
  \item A rotational period of 9 hours, with the rotational velocity vector distributed between the $\unit X$, $\unit Y$, and $\unit Z$ axes in a $1:2:-2$ ratio.
  \item An asteroid with radius $a_m = 1$ km and $K_{3m}=0$. For $K_{22}$ and $K_{20}$, we use two standard values. One with $(K_{22}, K_{20}) = (0, -0.097)$ and one with $(0.052, -0.202)$. Including the third point obtained by reflection $K_{22}\rightarrow -K_{22}$, these are the three points that minimize the mean distance between an arbitrary point in the allowed parameter space (equation \ref{eqn:parameter-bounds}) and these reference values. The first point is called the symmetric case because the corresponding uniform-density-ellipsoid model is rotationally symmetric around $\unit z$. The second case and its reflection are called the asymmetric cases. Values of $(0.052, -0.202)$ have $a < b$ in the ellipsoid model, and the reflected value has $a > b$. If not specified, we use the $a < b$ case. Specifically, the asymmetric case has $a=1140$ m, $b=1839$ m, and $c=565$ m, while the symmetric case has $a=b=1411$ m and $c=1008$ m.
\end{enumerate}
